---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mgi.20240527

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/patterninstitute/mgi.20240527/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/patterninstitute/mgi.20240527/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `{mgi.20240527}` package is an R data annotation tool designed to provide
[Mouse Genome Informatics (MGI)](https://www.informatics.jax.org/) gene marker
identifiers, symbols, and names. It offers essential functionality for quick
mapping between symbols, identifiers, and names, as well as converting to and
from Entrez and Ensembl gene identifiers, aligned with the latest gene models.
This package is based on the data release from May 27, 2024, ensuring users have
access to the most current information available for their genomic research
needs.

## Installation

``` r
# install.packages("pak")
pak::pak("patterninstitute/mgi.20240527")
```

## MGI gene markers

The list of gene markers returned by the function `markers()` corresponds
to those for which an associated gene model existed as of 27 May 2024.

```{r}
library(mgi.20240527)

markers()
```

## Crosswalk between identifiers or symbols

```{r}
# As of 27 May 2024, Dcdc3 symbol is deprecated in favor of Rp1.
update_symbol("Dcdc3")

# Because Dcdc3 is remapped to Rp1 both symbols are mapped to MGI:1341105.
symbol_to_marker_id("Dcdc3")
symbol_to_marker_id("Rp1")

symbol_to_marker_id("Dcdc3")
symbol_to_ensembl_id("Rp1")
```

## Markers with genomic coordinates

```{r}
markers("MGI:87860")

markers("ENSMUSG00000066583", "ensembl_gene_id")
```

## Old symbols' remapping

```{r}
symbols <- c(
  "Xkr4",
  "LOC102640625",
  "sid2057",
  "Sox17",
  "ENSMUSG00000074760",
  "R75157",
  "6430567E01Rik",
  "VINAS"
)

# Current MGI marker symbols as of 2024-05-27.
update_symbol(symbols)

# Map MGI marker symbols to marker identifiers.
symbol_to_marker_id(symbols)

# Map MGI marker symbols to Ensembl identifiers.
symbol_to_ensembl_id(symbols)

# Map MGI marker symbols to Entrez identifiers.
symbol_to_entrez_id(symbols)

# Deprecated Ensembl identifier, e.g. 1700010L13Rik (MGI:1922878) should map
# to ENSMUSG00000074760.
# URL: https://www.informatics.jax.org/marker/MGI:1922878
symbol_to_marker_id("1700010L13Rik")

# But results in NA because ENSMUSG00000074760 has status Retired:
# URL: https://www.ensembl.org/Mus_musculus/Gene/Idhistory?g=ENSMUSG00000074760
symbol_to_ensembl_id("1700010L13Rik")
```


## Disclaimer

The `{mgi.20240527}` package is provided "as is" without any warranties, express or implied. While we strive to ensure the accuracy and reliability of the data, we do not guarantee its completeness or correctness. The package is based on the data release from May 27, 2024, and may not include subsequent updates or corrections. Users are encouraged to verify the information independently and use it at their own risk. The authors and maintainers of this package are not liable for any direct, indirect, incidental, or consequential damages arising from its use.

## References

- Baldarelli RM, Smith CL, Ringwald M, Richardson JE, Bult CJ, Mouse Genome Informatics Group. 2024. Mouse Genome Informatics: an integrated knowledgebase system for the laboratory mouse. Genetics. 2024 May 7;227(1):iyae031.

- Baldarelli RM, Smith CM, Finger JH, Hayamizu TF, McCright IJ, Xu J, Shaw DR, Beal JS, Blodgett O, Campbell J, Corbani LE, Frost PJ, Giannatto SC, Miers DB, Kadin JA, Richardson JE, Ringwald M. 2021. The mouse Gene Expression Database (GXD): 2021 update. Nucleic Acids Res. 2021 Jan 8;49(D1):D924-D931.

- Krupke DM, Begley DA, Sundberg JP, Richardson JE, Neuhauser SB, Bult CJ. The Mouse Tumor Biology Database: A Comprehensive Resource for Mouse Models of Human Cancer. Cancer Res. 2017 Nov 1;77(21):e67-e70.
